---
title: "Exploratory DVs"
author: "Zeynep Enkavi"
output: 
html_document:
toc: true
toc_depts: 2
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(GGally)
library(plyr)
library(dplyr)
library(tidyr)
```

# Exploratory DVs

Since I am currently unable to successfully recover parameters on simulated data as described above here are some exploratory dependent variables that might capture the question of interest more simply.

To take a step back, this sample consists of three age groups: children, teens and adults and we hypothesize that the increase in risk taking behavior with age is related to sensitivity to high variance feedback. [[That is, children are too sensitive rare but large losses and therefore less likely to risk playing those machines again while teens and young adults weigh these losses more appropriately ...]]

Had I been able to fit the models correctly, the parameter of interest to observe the age differences on would have been an exponent on the prediction errors (as Sarah fit them or on rewards as we may have fitted them).

Here are some exploratory thoughts on other ways to quantify how sensitive subjects are to feedback from different machines.

```{r echo=FALSE}
##################
#Machine game data
##################
file_list <- list.files("/Users/zeynepenkavi/Downloads/machine_game")

for (file in file_list){
  
  tmp <- read.csv(paste0("/Users/zeynepenkavi/Downloads/machine_game/",file))
  tmp$Sub_id <- as.numeric(gsub("[^0-9]", "", file))
  tmp$Trial_number <- 1:nrow(tmp)
  
  if('X' %in% names(tmp)){
    tmp <- tmp[,-which(names(tmp) == "X")]
  }
  
  if (file == file_list[1]){
    machine_game_data = tmp
  }
  else{
    machine_game_data = rbind(machine_game_data, tmp)
  }
  rm(tmp)
}

rm(file, file_list)

#Remove subjects with incomplete data
incomplete_subs <- as.numeric(names(which(table(machine_game_data$Sub_id)!=180)))

machine_game_data_clean <- machine_game_data[machine_game_data$Sub_id %in% incomplete_subs == F,]

rm(incomplete_subs)

#Add cols for machine properties
assign.machine.info <- function(data){
  data$facet_labels <- with(data, ifelse(Trial_type == 1, "+5,-495", ifelse(Trial_type == 2, "-5,+495", ifelse(Trial_type == 3, "-10,+100", ifelse(Trial_type == 4, "+10,-100", NA)))))
  
  data$gain_mag <- with(data, ifelse(Trial_type == 1, "5", ifelse(Trial_type == 2, "495", ifelse(Trial_type == 3, "100", ifelse(Trial_type == 4, "10", NA)))))
  
  data$loss_mag <- with(data, ifelse(Trial_type == 1, "495", ifelse(Trial_type == 2, "5", ifelse(Trial_type == 3, "10", ifelse(Trial_type == 4, "100", NA)))))
  
  data$gain_freq <- with(data, ifelse(Trial_type == 1, "90", ifelse(Trial_type == 2, "10", ifelse(Trial_type == 3, "50", ifelse(Trial_type == 4, "50", NA)))))
  
  data$loss_freq <-  with(data, ifelse(Trial_type == 1, "10", ifelse(Trial_type == 2, "90", ifelse(Trial_type == 3, "50", ifelse(Trial_type == 4, "50", NA)))))
  
  data$magnitude <- with(data, ifelse(Trial_type == 1, "large", ifelse(Trial_type == 2, "large", ifelse(Trial_type == 3, "small", ifelse(Trial_type == 4, "small", NA)))))
  
  data$variance <- with(data, ifelse(Trial_type == 1, "high", ifelse(Trial_type == 2, "high", ifelse(Trial_type == 3, "low", ifelse(Trial_type == 4, "low", NA)))))
  
  return(data)
}

machine_game_data_clean <- assign.machine.info(machine_game_data_clean)

assign.age.info <- function(data){
  data$age_group <- with(data, ifelse(Sub_id<200000, "kids", ifelse(Sub_id>200000 & Sub_id<300000, "teen", "adult")))
  
  data %>% 
    group_by(Sub_id) %>%
    mutate(age = demog_data$calc_age[demog_data$subj_id])
}

##################
#Demographics data
##################
demog_data <- read.csv('/Users/zeynepenkavi/Dropbox/PoldrackLab/Developmental study/Final Redcap Data/DevelopmentalStudy_DATA_2015-03-25_1258.csv')

# head(names(demog_data), 20)
# str(demog_data$subj_id)
# str(demog_data$calc_age)

# Add age data to machine_game_data
machine_game_data_clean <- merge(machine_game_data_clean, demog_data[,c('subj_id', 'calc_age')], all.x = T, by.x = 'Sub_id', by.y = 'subj_id')

# Check if there are any problems with the merged ages
# summary(machine_game_data_clean$calc_age)
# which(is.na(machine_game_data_clean$calc_age))

##########
#Bart data
##########
file_list <- list.files("/Users/zeynepenkavi/Downloads/bart_tsv", pattern = '*.tsv')

for (file in file_list){
  
  tmp <- read.csv(paste0("/Users/zeynepenkavi/Downloads/bart_tsv/",file), sep = "\t")
  tmp$Sub_id <- as.numeric(strsplit(file, "_")[[1]][1])
  
  if (file == file_list[1]){
    bart_data = tmp
  }
  else{
    bart_data = rbind(bart_data, tmp)
  }
  rm(tmp)
}

rm(file, file_list)
```

## Average number of trials to play again post loss

One way to quantify how subjects respond to experiencing losses is to count how many trials it takes them to decide to play again. Here's a function that would calculate this.

```{r warning=FALSE}
count.postloss.trials <- function(subject_data){
  
  loss_trials = which(subject_data$Points_earned<0)
  
  play_trials= which(subject_data$Response == 1)
  
  post_loss_trials = play_trials[which(play_trials %in% loss_trials)+1]
  
  num_trials_post_loss = post_loss_trials - loss_trials
  
  return(data.frame(num_trials_post_loss = num_trials_post_loss))
}

# subject_data <- machine_game_data_clean[1:180,]
# tmp <- count.postloss.trials(subject_data)
```

This histogram suggests that ignoring which machine the loss is experienced on and which machine they play on next it takes subjects around two trials to play again after a loss. 

This seems to suggest that behavior almost doesn't change at all after a loss and that is (hopefully) unlikely to be true. So let's look at how many trials it takes a subject to play again on each machine after experiencing a loss on that machine.

```{r warning=FALSE}
postloss_trials <- machine_game_data_clean %>%
  group_by_(.dots = list(~Sub_id, ~Trial_type)) %>%
  do(count.postloss.trials(.))  %>%
  do(assign.machine.info(.))

mean_postloss_trials <- postloss_trials %>%
  group_by(Sub_id) %>%
  summarise(mean_post_loss = mean(num_trials_post_loss, na.rm=T))

mean_postloss_trials %>% ggplot(aes(mean_post_loss))+
  geom_histogram()+
  theme_bw()
```

It looks like it takes subjects fewer trials to play a machine again for machines with smaller losses compared to larger losses. For larger losses they are also faster to play it again if the losses are less frequent.

```{r warning=FALSE}
mean_postloss_trials_per_machine <- postloss_trials %>%
  group_by_(.dots = list(~Sub_id, ~Trial_type)) %>%
  summarise(mean_post_loss = mean(num_trials_post_loss, na.rm=T)) %>%
  do(assign.machine.info(.))

mean_postloss_trials_per_machine %>% ggplot(aes(mean_post_loss))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(~facet_labels)

mean_postloss_trials_per_machine %>% 
  group_by(facet_labels) %>%
  summarise(mean=mean(mean_post_loss, na.rm=T), sd = sd(mean_post_loss, na.rm=T))
```

Let's check if these distributions are different from one another with a paired t-test.

It seems that only the first panel is significantly different than the others. That is, subjects are faster to play the machine when they experience small losses half of the time compared to the other machines. 

```{r warning=FALSE}
with(mean_postloss_trials_per_machine, pairwise.t.test(mean_post_loss, facet_labels, p.adjust.method = 'bonferroni', paired=T))
```

What we are interested in is whether there is an age difference in each of these panels, or even more specifically whether there is an age difference in the last panel and whether this correlates with risk taking behavior (i.e. BART).

```{r warning=FALSE}
mean_postloss_trials_per_machine %>%
  mutate(age_group = ifelse(Sub_id<200000, "kids", ifelse(Sub_id>200000 & Sub_id<300000, "teen", "adult"))) %>%
  group_by_(.dots = list(~age_group,~facet_labels)) %>%
  summarise(age_group_mean = mean(mean_post_loss,na.rm=T),
            age_group_sem = )
```


```{r}
#Does this correlate with age or bart adjusted pumps? No.

adjusted.pumps <- function(subject_data){
  subject_data_adjusted = subject_data[subject_data$exploded == 0,]
  subject_pumps <- subject_data_adjusted %>% 
    group_by(trial.num) %>%
    summarise(total_pumps = sum(finished))
  out <- data.frame(Sub_id = unique(subject_data$Sub_id), mean_adjusted_pumps = mean(subject_pumps$total_pumps))
  return(out)
}

#subject_data <- bart_data[c(bart_data$Sub_id == 100003),]
#adjusted.pumps(subject_data)

bart_pumps <- ddply(bart_data, .(Sub_id), adjusted.pumps)

mean_postloss_trials <- merge(mean_postloss_trials, bart_pumps, all.x = T, by= 'Sub_id')

ggpairs(mean_postloss_trials, (which(names(mean_postloss_trials) %in% c('calc_age', 'mean_post_loss', 'mean_adjusted_pumps'))))
```

## Mean p(play) post loss
Operationalized as: Number of times you played immediately after experiencing a loss/Number of loss trials

```{r}
mean.postloss.play.prob <- function(subject_data){
  
  Sub_id = unique(subject_data$Sub_id)
  
  loss_trials = which(subject_data$Points_earned<0)
  
  mean_post_loss_prob <- mean(ifelse(subject_data$Response[loss_trials+1] == 2, 1, 0), na.rm=T)
  
  out <- data.frame(Sub_id = Sub_id, mean_post_loss_prob = mean_post_loss_prob)
  
  return(out)
}

# subject_data <- machine_game_data_clean[1:180,]
#tmp <- mean.postloss.play.prob(subject_data)

mean_postloss_trials <- merge(mean_postloss_trials, ddply(machine_game_data_clean, .(Sub_id), mean.postloss.play.prob), all.x=T, by='Sub_id')

ggpairs(mean_postloss_trials, (which(names(mean_postloss_trials) %in% c('calc_age', 'mean_post_loss', 'mean_adjusted_pumps', 'mean_post_loss_prob'))))
```

## How many people experience multiple losses (for each machine)

```{r}
loss.per.machine <- function(subject_data){
  
  Sub_id = unique(subject_data$Sub_id)
  
  loss_trials = which(subject_data$Points_earned<0)
  
  loss_per_machine = as.data.frame(table(subject_data$Trial_type[loss_trials]))
  
  out <- data.frame(Sub_id = rep(Sub_id, nrow(loss_per_machine)), machine = loss_per_machine$Var1, num_loss = loss_per_machine$Freq)
  
  return(out)
}

#subject_data <- machine_game_data_clean[1:180,]
#tmp <- loss.per.machine(subject_data)

loss_per_machine <- ddply(machine_game_data_clean, .(Sub_id), loss.per.machine)

loss_per_machine$facet_labels <- with(loss_per_machine, ifelse(machine == 1, "+5,-495", ifelse(machine == 2, "-5,+495", ifelse(machine == 3, "-10,+100", ifelse(machine == 4, "+10,-100", NA)))))

head(loss_per_machine)

loss_per_machine %>%
  ggplot(aes(num_loss))+
  geom_histogram()+
  facet_wrap(~facet_labels)+
  theme_bw()

#With machine do people experience a single loss most frequently?
with(loss_per_machine[loss_per_machine$num_loss == 1,], table(facet_labels))
```

## Cross-talk between machines
are you less likely to play overall after a loss or only less likely to play that machine

```{r}
mean_postloss_play_prob_per_machine <- ddply(machine_game_data_clean, .(Sub_id, Trial_type), mean.postloss.play.prob)

mean_postloss_play_prob_per_machine$facet_labels <- with(mean_postloss_play_prob_per_machine, ifelse(Trial_type == 1, "+5,-495", ifelse(Trial_type == 2, "-5,+495", ifelse(Trial_type == 3, "-10,+100", ifelse(Trial_type == 4, "+10,-100", NA)))))

mean_postloss_play_prob_per_machine %>% ggplot(aes(mean_post_loss_prob, fill = facet_labels))+
  geom_histogram(alpha = 0.5)+
  theme_bw()#+
#facet_wrap(~facet_labels)

mean_postloss_trials%>% ggplot(aes(mean_post_loss_prob))+
  geom_histogram()+
  theme_bw()

tmp <- mean_postloss_trials[,c("Sub_id", "mean_post_loss_prob")]
tmp$Trial_type <- 0
tmp$facet_labels <- 'Overall'
tmp <- tmp[,c(3,1,2,4)]
tmp <- rbind(tmp, mean_postloss_play_prob_per_machine)

tmp %>% ggplot(aes(mean_post_loss_prob, fill = facet_labels))+
  geom_histogram(alpha=0.5)+
  theme_bw()


pairwise.t.test(tmp$mean_post_loss_prob, tmp$facet_labels)

t.test(tmp$mean_post_loss_prob[tmp$facet_labels=='Overall'], tmp$mean_post_loss_prob[tmp$facet_labels=='-10,+100'])

t.test(tmp$mean_post_loss_prob[tmp$facet_labels=='Overall'], tmp$mean_post_loss_prob[tmp$facet_labels=='+10,-100'])

t.test(tmp$mean_post_loss_prob[tmp$facet_labels=='Overall'], tmp$mean_post_loss_prob[tmp$facet_labels=='-5,+495'])

t.test(tmp$mean_post_loss_prob[tmp$facet_labels=='Overall'], tmp$mean_post_loss_prob[tmp$facet_labels=='+5,-495'])

#Less likely to play after a loss trial in the low varince positive EV machine and more likely to play after a loss trial in the low variance positive EV machine

#What is optimal here? Is this pattern 'rational'?
#I don't think so. In the positive EV machine the optimal thing would be to always play. The decrease in probability in playing (from 0.5) after experiencing small losses would lead one to make less than one could
#Conversely in the low variance negative expected value machine one should never play. But the increase in probability in playing after losses 
```
