\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Neuroimage} % Sets Journal name


\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Insufficient and inappropriate model comparison can},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc citation processing

% Pandoc header



\begin{document}
\begin{frontmatter}

  \title{Insufficient and inappropriate model comparison can}
    \author[California Institute of Technology]{A. Zeynep Enkavi\corref{1}}
   \ead{zenkavi@caltech.edu} 
    \author[Stanford University]{Russell A. Poldrack}
   \ead{russpold@stanford.edu} 
      \address[California Institute of Technology]{Division of Humanities and Social Sciences, 1200 E. California Blvd.,
Pasadena, CA, 91125}
    \address[Stanford University]{Department of Psychology, 450 Jane Stanford Way, Stanford, CA, 94305}
      \cortext[1]{Corresponding Author}
  
  \begin{abstract}
  This is the abstract.
  
  It consists of two paragraphs.
  \end{abstract}
  
 \end{frontmatter}

\emph{Text based on elsarticle sample manuscript, see
\url{http://www.elsevier.com/author-schemas/latex-instructions\#elsarticle}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

\hypertarget{behavioral-data}{%
\subsection{Behavioral data}\label{behavioral-data}}

\hypertarget{outside-the-scanner}{%
\subsubsection{Outside the scanner}\label{outside-the-scanner}}

\hypertarget{inside-the-scanner}{%
\subsubsection{Inside the scanner}\label{inside-the-scanner}}

\hypertarget{neuroimaging-protocol}{%
\subsection{Neuroimaging protocol}\label{neuroimaging-protocol}}

\hypertarget{neuroimaging-preprocessing}{%
\subsection{Neuroimaging
preprocessing}\label{neuroimaging-preprocessing}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{behavioral-results}{%
\subsection{Behavioral results}\label{behavioral-results}}

\hypertarget{neuroimaging-results}{%
\subsection{Neuroimaging results}\label{neuroimaging-results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use BibTeX~to generate your bibliography and include DOIs
whenever available.

Here are two sample references: One in text Feynman and Vernon Jr.
(1963) and another at the sentence end (Dirac, 1953).

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Dirac1953888}{}%
Dirac, P.A.M., 1953. The lorentz transformation and absolute time.
Physica 19, 888--896.
doi:\href{https://doi.org/10.1016/S0031-8914(53)80099-6}{10.1016/S0031-8914(53)80099-6}

\leavevmode\hypertarget{ref-Feynman1963118}{}%
Feynman, R.P., Vernon Jr., F.L., 1963. The theory of a general quantum
system interacting with a linear dissipative system. Annals of Physics
24, 118--173.
doi:\href{https://doi.org/10.1016/0003-4916(63)90068-X}{10.1016/0003-4916(63)90068-X}


\end{document}


